\Title

### 2015-08-24 moho-inversion.tex
Finish parametrization. Formulate gauss-newton.

%\documentclass[]{article}
%\usepackage{natbib}
%\documentclass[extra,mreferee]{gji}
\documentclass[extra]{gji}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage[utf8]{inputenc}
\newcommand{\mbf}[1]{\mathbf{{#1}}}

@@

\usepackage[pdftex,colorlinks=true,hidelinks]{hyperref}
\hypersetup{
    pdftitle={\Title},
    pdfauthor={Leonardo Uieda (leo@leouieda.com)},
    pdfsubject={},
    pdfkeywords={\Keywords},
    pdfcreator={pdfTeX}
}

@@

@@

We parameterize the forward problem by discretizing the Moho undulation around
the reference depth into tesseroids.

Only Moho topography around a reference level.
\todo{Include figure of Moho with all else  removed.}
Discretize into tesseroids.
Parameters are the Moho depths.
Establish that the reference level ($h_{ref}$)  and density contrast
($\Delta\rho_j$) are hyper-parameters.

@@

Formulate regularized inversion using Gauss-Newton and Steepest Descent.
Use full Jacobian.
Steepest has poor convergence when close to the solution
\citep{kelley_iterative_1987}.

>>>>

%\documentclass[]{article}
%\usepackage{natbib}
%\documentclass[extra,mreferee]{gji}
\documentclass[extra]{gji}
\usepackage{timet}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}
\usepackage[utf8]{inputenc}

@@

\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
    pdftitle={\Title},
    pdfauthor={Leonardo Uieda (leo@leouieda.com)},
    pdfsubject={},
    pdfkeywords={\Keywords},
    pdfcreator={pdfTeX},
    allcolors=blue,
}

@@

%\newcommand{\mbf}[1]{\mathbf{{#1}}}
\newcommand{\mbf}[1]{\hbox{\mathversion{bold}$#1$}}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/paper/tesseroid-coord-sys}
    \caption{Sketch of a tesseroid (spherical prism) in a geocentric coordinate
        system (X, Y, Z).
        Observations are made at point P with respect to it's local
        North-oriented coordinate system (x, y, z).
        After \citet{uieda_tesserioid_2015},
        available at
        \url{http://dx.doi.org/10.6084/m9.figshare.1495525}.
    }
    \label{fig:tess}
\end{figure}

We parameterize the forward problem by discretizing the Moho undulation
into a grid of $M_{lon} \times M_{lat} = M$ juxtaposed tesseroids (\fig{tess}).
\todo{Cite source separation figure}
In cases where the Moho is above the reference depth ($z_{ref}$),
the top of the tesseroid is the Moho depth at the location of the $k$-eth
tesseroid ($z_{k}$),
the bottom is $z_{ref}$, and the density-contrast is positive.
If the Moho is below the $z_{ref}$, the top of the tesseroid is $z_{ref}$,
the bottom is the Moho depth, and the density-contrast is negative.

Considering that the density-contrast of each tesseroid is fixed,
the predicted gravity anomaly of the Moho is a non-linear function of the
parameters $z_k$, $k=1, \ldots, M$,

\begin{equation}
    d_i = f_i(\mbf{p}),
\end{equation}

\noindent in which $d_i$ is the $i$th element of the predicted data vector
$\mbf{d}$, $\mbf{p}$ is the parameter vector containing the Moho depths
($z_k$),
and
$f_i$ is the $i$th non-linear function that maps the parameters onto the data.

@@

We wish to estimate the parameter vector $\mbf{p}$ from a set of observed
gravity anomaly data $\mbf{d}^o$.
The least-squares estimate is the one that minimizes the data-misfit function

\begin{equation}
    \phi(\mbf{p}) = \|\mbf{d}^o - \mbf{d}(\mbf{p})\|_2^2
    = [\mbf{d}^o - \mbf{d}(\mbf{p})]^T[\mbf{d}^o - \mbf{d}(\mbf{p})].
\end{equation}

\noindent However, non-linear inversions for the relief of an interface (like
the Moho) are ill-posed and required additional constraints in the form of
regularization \citep{silva_potential-field_2001}.
We will use the first-order Tikhonov regularization to impose smoothness on the
solution,

\begin{equation}
    \theta(\mbf{p}) = \|\mbf{R}\mbf{p}\|^2_2
    = \mbf{p}^T\mbf{R}^T\mbf{R}\mbf{p},
\end{equation}

\noindent where $\mbf{R}$ is an $L \times M$ finite-difference matrix
representing the $L$ first-order differences between adjacent tesseroids.

The solution $\hat{\mbf{p}}$ to the regularized inverse problem is the one that
minimizes the goal function

\begin{equation}
    \Gamma(\mbf{p}) = \phi(\mbf{p}) + \mu\theta(\mbf{p}).
\end{equation}

\noindent $\Gamma(\mbf{p})$ is non-linear with respect to $\mbf{p}$.
Thus, we can find the minimum of the goal function using gradient-based
iterative optimization
methods like the Gauss-Newton method or the Steepest Descent method.
Such methods start from an initial estimate $\mbf{p}^0$ and  repeatedly update
the estimate.
The update at the $k$th iteration $\mbf{\Delta p} = \mbf{p}^{k+1} - \mbf{p}^k$
for the Gauss-Newton method is

\begin{equation}
    \mbf{H}(\mbf{p}^k)\mbf{\Delta p} = -\mbf{\nabla}\Gamma(\mbf{p}^k),
\end{equation}

\noindent in which
$\mbf{H}$ is the Hessian matrix of $\Gamma$,

\begin{equation}
    \mbf{H}(\mbf{p}^k) = 2\mbf{A}(\mbf{p}^k)^T\mbf{A}(\mbf{p}^k) +
    2\mu\mbf{R}^T\mbf{R},
\end{equation}

\noindent
$\mbf{\nabla}\Gamma$ is the gradient vector of $\Gamma$,

\begin{equation}
    \mbf{\nabla}\Gamma(\mbf{p}^k) = -2\mbf{A}(\mbf{p}^k)^T[\mbf{d}^o - \mbf{d}] +
    2\mu\mbf{R}^T\mbf{R}\mbf{p}^k,
\end{equation}

\noindent and
$\mbf{A}$ is the Jacobian or sensitivity matrix,

\begin{equation}
    A_{ij} = \dfrac{\partial f_i}{\partial p_j}(\mbf{p}).
\end{equation}

Steepest has poor convergence when close to the solution
\citep{kelley_iterative_1987}.
